Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-20 ba_resubmit/deprecated/kundinger_resubmit3.tex
cleaning up repo.

Consider two data files $\bm{X}_1$ and $\bm{X}_2$ containing records $\{x_{1i}\}_{i=1}^{n_1}$ and $\{x_{2j}\}_{j=1}^{n_2}$ respectively. Without loss of generality, denote files such that $n_1 \geq n_2$. For ease of readability, we follow the convention established by \cite{sadinle_bayesian_2017} and say ``record $i \in \bm{X}_1$" rather than the more compact $x_{1i}$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_1 \times n_2}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_2})$ for the records in $\bm{X}_2$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  i \in \bm{X}_1 \; \text{and}\; j\in \bm{X}_2 \; \text{refer to the same entity}; \\
		n_1 + j, & \text{if record}\;  j \in \bm{X}_2 \; \text{does not have a match in}\; \bm{X}_1. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that $\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ We refer to set of all comparison vectors $\bm{\gamma}_{ij}$ as $\Gamma$.

%Frequently, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two data files.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for record $i$ and $j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for record $i$ and $j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for record $i$ and $j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

In the construction of comparison vectors, it is common to encounter missing information in record $i \in \bm{X}_1$ or $j \in \bm{X}_2$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs at random (MAR, per \cite{LittleRubin2002}). With the MAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise.

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999, jaro1989, larsen_2001, enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	& \Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

%\textcolor{red}{Each record pair is then independently classified as a match or nonmatch according to the} likelihood ratio  
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}.
%\end{align}
%This ratio is large when there is strong evidence of the pair being a match, and small otherwise. All record pairs with weights above a specified threshold are considered a match.

@@

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_1} \sum_{j=1}^{n_2} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_1}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_2; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_2}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_1. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

%The comparison data $\bm{\gamma}_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
%\end{align}
%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
%
%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
%
%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
%
%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework. In addition to standard Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$. He assigns a prior distribution for the probability of the indicator that a record in $\bm{X}_2$ has a match in $\bm{X}_1$, so that $I(Z_j \leq n_1) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$. It follows that the number of records in $\bm{X}_2$ that have matches, denoted $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$, is distributed according to a $\text{Beta-Binomial}(n_2, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $\bm{X}_2$ that have matches, formally denoted $\{I(Z_j \leq n_1)\}_{j=1}^{n_2}$, all $n_1 ! / (n_1 - n_{12}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the beta distribution for bipartite matching, is given by
\begin{align}
\label{eqn:sadinle_prior}
P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each \(Z_j\). This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

%hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X}_2$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the number of records in $\bm{X}_2$ that have a match in $\bm{X}_1$. 
%One criticism of the independent matching assumption is that the decision rule leads to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%\begin{align}
%	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m}),  \\
%	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag, \\
%	& \bm{Z} \sim \mathcal{B}, \notag
%\end{align}
%where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m})$ and $\mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the beta distribution for bipartite matching, given by
%$$P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\bm{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\bm{Z}) + \alpha_{\pi}, n_2 - n_{12}(\bm{Z}) - \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
%$$
%where $\text{B}(\cdot, \cdot)$ represents the Beta function, hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$, and $n_{12}(\bm{Z}) = \sum_{j=1}^{n_2} I(Z_j \leq n_1 + j)$ is the records in $\bm{X}_2$ that have a match in $\bm{X}_1$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate records when sampling \(Z_j\). This creates a dependency that makes the sampler inherently serial, which can slow down the sampler when working on large linkage tasks.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use a proper prior. We denote the fast Beta prior as follows. For each $Z_j$, we have
\begin{align}
	\label{eqn:fast_beta_prior}
	p(Z_j = z_j| \pi) &= \begin{cases} 
	\frac{1}{n_1}\pi,  & z_j \leq n_1; \\
	1-\pi, &  z_j  = n_1 + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and each record $i \in \bm{X}_1$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $\bm{X}_2$ that have matches in $\bm{X}_1.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_1, n_1 +j\}$ is a priori equally likely for $Z_j$. This however amounts to a prior probability of $n_1 / (n_1 + 1)$ that record $j$ has a match in $\bm{X}_1$. In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_1 \right)$ random variable.

%Intuitively, this set of priors says that record $j \in \bm{X}_2$ has some match in $\bm{X}_1$ with probability $\pi$, and that each record $i \in \bm{X}_1$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. One non-informative choice might be \(\pi \sim \text{Beta}(1, 1)\), which corresponds to a prior belief that non-matches and matches are equally likely, and another might be \(\pi \sim \text{Beta}\left(1, \frac{1}{n_1}\right)\), which corresponds to a uniform prior on the labeling of \(\bm{Z}\). We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_1}\right)$ random variable.

Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $\bm{X_2}$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches. We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case. In this setting, there are at most $n_2$ matching pairs out of $n_1 n_2$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_1}$ and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi-Sunter model makes $n_1 \times n_2$ independent matching decisions and \texttt{BRL} makes $n_2$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_2$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.

%To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}, reproduced in Appendix \ref{bayes-estimate}. Since the prior does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. In the event that we have two records $j$ and $j'$ such that both $P(\hat{Z}_j = i |\Gamma) > \frac{1}{2}$ and $ P(\hat{Z}_{j'} = i |\Gamma) > \frac{1}{2}$, we accept the match with the higher posterior probability, and declare the other to have no match. A similar approach appears in the most probable maximal matching sets used by \cite{steorts_bayesian_2016} to match records to latent entities.

>>>>

%Consider two data files $A$ and $B$ containing records $\{x_{1i}\}_{i=1}^{n_A}$ and $\{x_{2j}\}_{j=1}^{n_B}$ respectively.

Consider two data files $A$ and $B$, consisting of records $\{A_i: i = 1, \ldots, n_A\}$ and $\{B_j: j = 1, \ldots, n_B\}$. Suppose the files have $F$ fields in common to be used for linkage, and that these fields take on values $A_{if}$ and $B_{jf}$ respectively. Without loss of generality, denote files such that $n_A \geq n_B$. Under bipartite matching, the set of matches across data files can be represented in two equivalent ways. First, we may use a matrix $\Delta \in \{0, 1\}^{n_A \times n_B}$, where
\begin{align}
	\Delta_{ij} =
	\begin{cases}
		1, & \text{if records}\;  A_i \; \text{and}\; B_j \; \text{refer to the same entity}; \\
		0, & \text{otherwise}.\\
	\end{cases}
\end{align}
This sparse matrix representation can become cumbersome for large linkage tasks. More compactly, bipartite matching also can be viewed as a labeling $\bm{Z} = (Z_1, \ldots, Z_{n_B})$ for the records in $B$ such that 
\begin{align}
	Z_{j} =
	\begin{cases}
		i, & \text{if records}\;  A_i \; \text{and}\; B_j  \; \text{refer to the same entity}; \\
		n_A + j, & \text{if record}\;  B_j \; \text{does not have a match in}\; A. \\
	\end{cases}
\end{align}
We can go back and forth between the two using $\Delta_{ij} = I(Z_j = i),$ where $I(\cdot) = 1$ when the expression inside the parentheses is true, and $I(\cdot) = 0$ otherwise.

%{\color{red}{The following paragraph is pretty unnecessary, I never use this notation}}
Denote the set of matches by $\bm{M} = \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 1\}$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \leq n_A, j \leq n_B, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'' and to record pairs that are estimated as non-matches as ``non-links.''

@@

%{\color{red}{Beka suggested eliminating the following two paragraphs. Thoughts? I think the first one is good, the second seems unneccesary.}}

Intuitively, matching records (those that refer to the same entity) should be similar; records that are non-matching should be dissimilar. \cite{fellegi_theory_1969} proposed encoding this using a comparison vector $\gamma_{ij}$ computed for each record pair $(i,j)$ in $A \times B.$ Denote the number of criteria for comparing records by $F$, such that $\gamma_{ij} = (\gamma_{ij}^1, \ldots, \gamma_{ij}^F).$ {\color{red}{We use $\Gamma_{ij}$ when referring to the vector as a random variable, and $\gamma_{ij}$ when referring to observed data. We collect the vectors $\gamma_{ij}$ as the comparison matrix $\gamma \in \mathbb{R}^{n_A n_B \times F}$}}.

%Frequently, $\gamma_{ij}$ consists of one comparison for each feature shared between the two data files.

The simplest way to compare any particular feature for two records is to check for exact agreement, and this is commonly used for categorical features. For example, if zip code is linking feature $f$, we can set $\gamma_{ij}^f=1$ when the zip codes for records $A_i$ and $B_j$ agree exactly, and set $\gamma_{ij}^f=2$ when they do not. For numerical features, we can use absolute difference between the two feature values. For example, if age is linking field $f$, we can set $\gamma_{ij}^f = 1$ when the ages for records $A_i$ and $B_j$ match exactly, $\gamma_{ij}^f = 2$ when the ages for records $A_i$ and $B_j$ are within one year but not equal, and $\gamma_{ij}^f = 3$ when the ages are two or more years apart. For text features, such as names, we can use string distance metrics such as Levenstein or Jaro-Winkler distance \citep{cohen2003comparison}. We then set thresholds that allow us to represent comparisons through discrete levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}.

More generally, let $\mathcal{S}_f(i,j)$ denote a similarity measure for feature $f$ of records $A_i$ and $B_j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f$ intervals denoted by $I_{f1}, \ldots, I_{fL_f}$. Here, $I_{f1}$ represents the highest level of agreement (including complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors such that $\gamma_{ij}^f = l \; \text{if} \; \mathcal{S}_f(i,j) \in I_{fl}.$ The choices of $I_{fl}$ are application specific, as we discuss in the simulation and case studies.

{\color{red}{
In the construction of comparison vectors, it is common to encounter missing information in record $A_i$ or $B_j$. As a result, the comparison vector $\mathbf{\gamma}_{ij}$ will have missing values. We assume that this missingness occurs completely at random (MCAR, per \cite{LittleRubin2002}). To notate a missing value in any $\gamma_{ij}^f$, we use $I_{obs}(\gamma_{ij}^f)=1$ when $\gamma_{ij}^f$ is observed and $I_{obs}(\gamma_{ij}^f)=0$ otherwise. With the MCAR assumption, we can marginalize over the missing data, and do all computation simply using the observed data. }
}

@@

The seminal \cite{fellegi_theory_1969} model employs two independence assumptions: first, that comparison vectors are conditionally independent given their matching status, and second, that the matching status of the record pairs are independent. Using these assumptions, \cite{winkler_state_1999}, \cite{jaro1989}, \cite{larsen_2001}, \cite{enamorado2019using} and others model the comparison data through mixture models of the form
\begin{subequations}
\begin{align}
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 1 \stackrel{iid}{\sim} \mathcal{M}(\bm{m}), \label{eqn:fs_model} \\
	&\Gamma_{ij} = \gamma_{ij} \mid \Delta_{ij} = 0  \stackrel{iid}{\sim} \mathcal{U}(\bm{u}), \label{eqn:m_dist}\\
	&\Delta_{ij}   \stackrel{iid}{\sim} \text{Bernoulli}(\lambda). \label{eqn:u_dist}
\end{align}
\end{subequations}
Here, $\mathcal{M}$ and $\mathcal{U}$ are the distributions for matching and non-matching record pairs, $\bm{m}$ and $\bm{u}$ are their respective sets of parameters, and $\lambda$ is the marginal probability that a record pair is a match. When using comparison vectors with discrete agreement levels, $\mathcal{M}$ and $\mathcal{U}$ are collections of independent multinomial distributions for each linkage feature. Accordingly, $\bm{m} = (\bm{m}_1, \ldots, \bm{m}_F)$, where $\bm{m}_f = (m_{f1}, \ldots, m_{fL_f})$ and $m_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 1)$ for all fields $f$ and agreement levels $l$. The $\bm{u}$ parameters are defined similarly, with $u_{fl} = P(\gamma_{ij}^f = l|\Delta_{ij} = 0)$.

@@

%To address this issue, \cite{jaro1989} proposed solving the following linear sum assignment problem:
%\begin{align}
%	\label{eqn:jaro}
%	&\max_{\Delta} \sum_{i=1}^{n_A} \sum_{j=1}^{n_B} w_{ij} \Delta_{ij} 
%	\quad \text{subject to} \quad \Delta_{ij} \in \{0,1\}; \\
%	& \quad \quad \sum_{i=1}^{n_A}  \Delta_{ij}  \leq 1, j=1,2, \ldots n_B; \; \text{and} \notag\\
%	& \quad \quad\sum_{i=1}^{n_B}  \Delta_{ij}  \leq 1, i=1,2, \ldots n_A. \notag
%\end{align}
%
%The solution to this problem, for which several simple algorithms exist, is a bipartite matching that maximizes the sum of the Fellegi-Sunter weights among matched pairs. Jaro provided no theoretical justification for using this approach; however \cite{sadinle_bayesian_2017} recently showed that under certain conditions, (\ref{eqn:jaro}) is the maximum likelihood estimate for bipartite matching.

%The comparison data $\gamma_{ij}$ are not sufficient to determine whether a record is a match or non-match because of errors that naturally occur in the data. This motivated \cite{fellegi_theory_1969} to consider the likelihood ratio
%\begin{align}
%	\label{eqn:wts}
%	w_{ij} = \frac{P(\gamma_{ij} \mid \Delta_{ij} = 1)}{P(\gamma_{ij} \mid \Delta_{ij} = 0)}
%\end{align}
%as a weight to estimate if record pairs are a match or non-match. This ratio will be large when there is strong evidence of the pair being a match, and small otherwise. 
%
%In practice, $P(\cdot \mid \Delta_{ij} = 1)$ and $P(\cdot \mid \Delta_{ij} = 0)$ are unknown to the modeler, and thus must be estimated. 
%
%%is the set of parameters $u_{fl} = P(\gamma_{ij}^f = l|Z_j \neq i)$ for the non-matching pairs.
%
%After estimating $\bm{m}$ and $\bm{u}$, the Fellegi-Sunter method uses thresholds $T_m$ and $T_u$ such that all record pairs with $w_{ij} > T_m$ are declared as links, and all those with $w_{ij} < T_u$ are declared as non-links (with those such that $T_u < w_{ij} < T_m$ left undetermined and subject to clerical review). This is a problem in practice as transitive closures can be violated. Since each $w_{ij}$ is calculated independently, there can be situations in which both $w_{ij} > T_m$ and $w_{i'j} > T_m$, so both $(i,j)$ and $(i, j')$ are declared as links, even though such a matching is not allowed by assumption.

@@

To estimate a one-to-one matching without using a post-processing step, \cite{sadinle_bayesian_2017} incorporates this constraint directly into the model through a Bayesian framework. In addition to {\color{red}{uniform}} Dirichlet priors for the $\bm{m}_f$ and $\bm{u}_f$ parameters, he proposes the ``beta distribution for bipartite matching" for the linkage parameter $\bm{Z}$. He assigns a prior distribution for the probability of the indicator that a record in $B$ has a match in $A$, so that $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$, where $\pi$ itself is taken to be distributed $\text{Beta}(\alpha_{\pi}, \beta_{\pi})$. It follows that the number of records in $B$ that have matches, denoted $n_{AB}(\bm{Z}) = \sum_{j=1}^{n_B} I(Z_j \leq n_A)$, is distributed according to a $\text{Beta-Binomial}(n_B, \alpha_{\pi}, \beta_{\pi})$. Conditioning on the set of records in $B$ that have matches, formally denoted $\{I(Z_j \leq n_A)\}_{j=1}^{n_B}$, all $n_A ! / (n_A - n_{AB}(\bm{Z}))!$ bipartite matchings are taken to be equally likely. Thus, the prior is given by
\begin{align}
\label{eqn:sadinle_prior}
P(\bm{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_A - n_{AB}(\bm{Z}))!}{n_A !}\frac{\text{B}(n_{AB}(\bm{Z}) + \alpha_{\pi}, n_B - n_{AB}(\bm{Z}) + \beta_{\pi})}{\text{B}(\alpha_{\pi}, \beta_{\pi})},
\end{align}
where $\text{B}(\cdot, \cdot)$ represents the Beta function. This prior strictly enforces one-to-one matching, inducing a Gibbs sampler that removes previously matched records from the set of candidate records when sampling each $Z_j$. This makes the sampler inherently serial, which can be slow when working on linkage tasks with more than a few thousand records.

@@

In contrast to the prior over the vector $\bm{Z}$ from \cite{sadinle_bayesian_2017}, we follow \cite{wortman2019} and use independent priors for each component $Z_j$. However, unlike \cite{wortman2019} who proposes a flat prior for $Z_j$, we use the fast Beta prior as follows. For each $Z_j$, we have
\begin{align}
	\label{eqn:fast_beta_prior}
	p(Z_j = q| \pi) &= \begin{cases} 
	\frac{1}{n_A}\pi,  & q \leq n_A; \\
	1-\pi, &  q  = n_A + j; \\
\end{cases} \\
\pi &\sim \text{Beta}(\alpha_{\pi}, \beta_{\pi}) \notag.
\end{align}
We can interpret (\ref{eqn:fast_beta_prior}) as follows: record $B_j$ has some match in $A$ with probability $\pi$, and each record $A_i$ is equally likely to be that match. The hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about  the proportion of records in $B$ that have matches in $A.$

In the \cite{wortman2019} flat prior, each value $\{1, \ldots, n_A, n_A +j\}$ is a priori equally likely for $Z_j$. This however amounts to a prior probability of $n_A / (n_A + 1)$ that record $B_j$ has a match in $A$. In our preliminary studies, the flat prior results in poor precision; hence, we prefer (\ref{eqn:fast_beta_prior}). We also note that the flat prior is equivalent to a special case of the fast Beta prior with $\pi$ fixed at the mean of a $\text{Beta}\left(1, 1 / n_A \right)$ random variable.

%Intuitively, this set of priors says that record $B_j$ has some match in $A$ with probability $\pi$, and that each record $A_i$ is equally likely to be that match. $\pi$ itself is given a prior distribution with hyperparameters $\alpha_{\pi}$ and $\beta_{\pi}$ to encode prior beliefs about  the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. One non-informative choice might be $\pi \sim \text{Beta}(1, 1)$, which corresponds to a prior belief that non-matches and matches are equally likely, and another might be $\pi \sim \text{Beta}\left(1, \frac{1}{n_A}\right)$, which corresponds to a uniform prior on the labeling of $\bm{Z}$. We note here that the prior from \cite{wortman2019} can be viewed as a special case of the fast beta prior, with $\pi$ fixed at the mean of a $\text{Beta}\left(1, \frac{1}{n_A}\right)$ random variable.

Note that linkage with \texttt{fabl} is conducted at the record level, rather than at the record pair level, as in the Fellegi-Sunter model. That is, $\pi$ under \texttt{fabl} estimates the proportion of records in $B$ that have matches, whereas $\lambda$ in the Fellegi-Sunter model estimates the proportion of record pairs that are matches. We find $\pi$ to be more a interpretable parameter than $\lambda$ in the bipartite case. In this setting, there are at most $n_B$ matching pairs out of $n_A n_B$ total pairs, meaning that $\lambda$ is bounded above by $\frac{1}{n_A}$ and tends towards 0 as the size of the linkage task grows. Additionally, while the Fellegi-Sunter model makes $n_A \times n_B$ independent matching decisions and \texttt{BRL} makes $n_B$ dependent matching decisions, \texttt{fabl} strikes a middle ground between the two, making $n_B$  independent matching decisions. As shown in Sections \ref{sec:simulations} and \ref{sec:case-studies}, this allows \texttt{fabl} to fit a Bayesian record linkage model like \texttt{BRL} while making computational efficiency gains possible by exploiting independence.

To obtain an estimate $\hat{\bm{Z}}$ of the linkage structure, we use the loss functions and Bayes estimate from \cite{sadinle_bayesian_2017}. However, there is a crucial difference: since \ref{eqn:fast_beta_prior} does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $B$ to one record in $A$. To achieve a Bayes estimate that fulfills the one-to-one matching requirement, we minimize the expected loss subject to the constraint that $\hat{Z}_j \neq \hat{Z}_{j'}$ for all $j \neq j'$. Details for the initial Bayes estimate and the post-processing procedure are provided in Appendix \ref{bayes-estimate}.

\begin{table}[t!]
	\centering
	\begin{tabular}[t!]{ll}
		Symbol & Description \\
		\hline
		$A, B$ & data files\\
		$i \in 1, \ldots, n_A $ & index over records in $A$\\
		$j \in 1, \ldots, n_B $ & index over records in $B$\\
		$f \in 1, \ldots F$ & index over fields used for comparisons \\
		$l \in 1, \ldots L_f$ & index over agreement levels for feature $f$ \\
		$n_{AB}$ & number of entities in common between $A$ and $B$\\
		$\gamma_{ij}$ & comparison vector for records $A_i$ and $B_j$ \\
		$Z_j = i$ & records $A_i$ and $B_j$ match \\
		$Z_j = n_A + j$ & record $B_j$ has no match in $A$ \\
		$m_{fl}$ & $P(\gamma_{ij}^f = l | Z_j = i)$ \\
		$u_{fl}$ & $P(\gamma_{ij}^f = l | Z_j \neq i)$ \\
		$\pi$ & probability that a record in $B$ has a match in $A$ \\
		\hline
	\end{tabular}\caption{Summary of model notation.}\label{table_notation_A}
\end{table}
